---
title: "Authentification of H. forskali"
author: "Frank DAVID"
date: '2026-01-15'
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
Sys.setlocale("LC_TIME", "C")
```

### R Markdown

This R Markdown document is made to support the findings of the paper "A screening of tools to differentiate wild vs. farmed sea cucumbers in multi-trophic aquaculture systems with no artificial feed inputs" by David et al. (2026). 


## Preparing the toolbox

Let's begin by loading the libraries that we will require to run the analyses.

```{r echo=TRUE, message=FALSE}
#library(agricolae)
#library(vegan)
library(ade4)
library(factoextra)
library(easyCODA)
library(pls)
library(rpart)

```


## Loading the data
We load fatty acids, stable isotopes, amino acids and sterols data from the GitHub server.

For details on line and column headings please check the readme file.

```{r}

FA <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-authenticity/main/FA_screening.csv", sep=";", header=T)
SI <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-authenticity/main/SI_screening.csv", sep=";", header=T)
AA <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-authenticity/main/AA_screening.csv", sep=";", header=T)
St <- read.csv("https://raw.githubusercontent.com/DAVID-Fk/Sea-cucumber-authenticity/main/St_screening.csv", sep=";", header=T)

```

## Fatty acids

```{r}
allFA=FA[,-c(1:7, which(colnames(FA)=="C23.0"))]
rownames(allFA)=FA[,1]
allFA[is.na(allFA)]=0

## Percentage table
prct=allFA/rowSums(allFA)*100

## Concentrations table 
conc=allFA/FA[,which(colnames(FA)=="C23.0")]*FA[,which(colnames(FA)=="StdC23")]/FA[,which(colnames(FA)=="SplMass")]
conctot=rowSums(conc)

# Let's define categories of FA

brFA=rowSums(prct[, c(which(colnames(prct)=="C15.0iso"), which(colnames(prct)=="C15.0anteiso"), which(colnames(prct)=="C16.0iso"))])
SFA=rowSums(prct[, c(which(colnames(prct)=="C14.0"), which(colnames(prct)=="C15.0"), which(colnames(prct)=="C16.0"), which(colnames(prct)=="C18.0"), which(colnames(prct)=="C19.0"),which(colnames(prct)=="C20.0"), which(colnames(prct)=="C21.0"), which(colnames(prct)=="C22.0"))])
PUFA=rowSums(prct[, c(which(colnames(prct)=="C18.2w6"), which(colnames(prct)=="C18.3w3"), which(colnames(prct)=="C18.4w3"), which(colnames(prct)=="C20.2w6"), which(colnames(prct)=="C20.4w6"), which(colnames(prct)=="C20.5w3"), which(colnames(prct)=="C22.5w6"), which(colnames(prct)=="C22.6w3"))])
HUFA=rowSums(prct[, c(which(colnames(prct)=="C20.4w6"), which(colnames(prct)=="C20.3w6"), which(colnames(prct)=="C20.5w3"), which(colnames(prct)=="C22.5w6"), which(colnames(prct)=="C22.6w3"))])
w3=rowSums(prct[, c(which(colnames(prct)=="C18.3w3"), which(colnames(prct)=="C18.4w3"), which(colnames(prct)=="C20.5w3"), which(colnames(prct)=="C22.6w3"))])
w6=rowSums(prct[, c(which(colnames(prct)=="C18.2w6"), which(colnames(prct)=="C20.3w6"), which(colnames(prct)=="C20.2w6"), which(colnames(prct)=="C20.4w6"), which(colnames(prct)=="C22.5w6"))])
LCMUFA=rowSums(prct[, c(which(colnames(prct)=="C20.1w9"), which(colnames(prct)=="C20.1w7"), which(colnames(prct)=="C22.1w11"), which(colnames(prct)=="C22.1w9"), which(colnames(prct)=="C23.1w9"), which(colnames(prct)=="C24.1w9"))])
MUFA=rowSums(prct[, c(which(colnames(prct)=="C16.1w7"), which(colnames(prct)=="C18.1w9"), which(colnames(prct)=="C18.1w7"), which(colnames(prct)=="C20.1w9"), which(colnames(prct)=="C20.1w7"), which(colnames(prct)=="C22.1w11"), which(colnames(prct)=="C22.1w9"), which(colnames(prct)=="C23.1w9"), which(colnames(prct)=="C24.1w9"))])

rowSums(cbind(brFA, SFA, PUFA, MUFA))

# Let's edit a summary table

summaryFA=data.frame(M=aggregate(cbind(prct, brFA, SFA, MUFA, LCMUFA, PUFA, HUFA, w3, w6, conctot) , by=list(FA$SplBatch), mean), SD=aggregate(cbind(prct, brFA, SFA, MUFA, LCMUFA, PUFA, HUFA, w3, w6, conctot), by=list(FA$SplBatch), sd), LE=aggregate(cbind(prct, brFA, SFA, MUFA, LCMUFA, PUFA, HUFA, w3, w6, conctot), by=list(FA$SplBatch), length))

#write.table(t(summaryFA), file="meansFA.txt")

```


### Graphical representation using PCA

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

#svg(file="Fig2A.svg", width=7.5, height=6)

par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

acp=dudi.pca(prct, scannf = FALSE, nf = 2)
varPCA=get_pca_var(acp)
varPCAhigh=which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*1.5|varPCA$contrib[,2]>100/nrow(varPCA$contrib)*1.5)
varPCAlow=which(varPCA$contrib[,1]<100/nrow(varPCA$contrib)*1.5&varPCA$contrib[,2]<100/nrow(varPCA$contrib)*1.5)

plot(acp$li[,c(1,2)], acp$co[,c(1,2)], type="n", ann=F, bty="n", axes=F, ylim=c(-7, 7), xlim=c(-10, 8))
axis(1, pos=0, at=c(seq(-10, 8, 2)[-6]), labels=F, tck=0.01, padj=-4)
axis(2, pos=0, at=c(seq(-6, 6, 2)[-4]), labels=F, tck=0.01, hadj=0.1)

title(xlab=paste("PC1 (", round((acp$eig/sum(acp$eig)*100)[1], 1), "%)", sep=""), line=0)
title(ylab=paste("PC2 (", round((acp$eig/sum(acp$eig)*100)[2], 1), "%)", sep=""), line=0)

arrows(x0=0, y0=0, x1=acp$co[,1][varPCAhigh]*8, y1=acp$co[,2][varPCAhigh]*8, length=0.1, col="grey40", lty=5)
text(acp$co[varPCAhigh,]*8.5, labels=colnames(prct)[varPCAhigh], col="grey40")

arrows(x0=0, y0=0, x1=acp$co[,1][varPCAlow]*8, y1=acp$co[,2][varPCAlow]*8, length=0.1, col="grey90", lty=5)
text(acp$co[varPCAlow,]*8.5, labels=colnames(prct)[varPCAlow], col="grey90")

a=ordihull(acp$li, groups=paste(FA$SplBatch, FA$Group), lty=4, label=F, display=c("sites"), draw="lines")

points(acp$li[, 1], acp$li[, 2], pch=23, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(FA$Group)], cex=1.5)

a=ordihull(acp$li, groups=paste(FA$SplBatch, FA$Group), border=NA, label=F, display=c("sites"), draw="polygon", col="white", alpha=50)

centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  

points(centroid, pch=21, cex=3.5, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1.2, font=2, col=c("white", "black", "black")[as.factor(var[,2])])

legend("topright" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1)

legend("bottomleft", legend=paste("N =", length(levels(as.factor(FA$SplBatch))),"; n =", nrow(acp$li)), bty="n", pt.cex=2)

legend("topleft", legend=expression(bold("A")), bty="n", cex=3, inset=c(-0.1,-0.05))


#dev.off()
  
```


### Graphical representation using PLSDA

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

# Transforming group into indicating variable
var.ind <- DUMMY(paste(FA$Group, sep=""))
fac <- as.factor(paste(FA$Group, sep=""))

# Put the matrix into the good "matrix" shaping
DATA.M <- as.matrix(prct)

# Canonical Powered Partial least squared (PLS) regression
PLSDA<-cppls(var.ind ~ DATA.M, ncomp=2, scale=T) 
summary (PLSDA)
attributes(PLSDA)
PLSDA$scores
att=attributes(loadings(PLSDA))$explvar

# Drawing

#svg(file="Fig2B_FA.PLSDA.svg", width=7.5, height=6)

par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

plot(PLSDA$scores, type="n", ann=F, bty="n", axes=F, xlim=c(-5, 8), ylim=c(-7, 6))

axis(1, pos=0, at=c(seq(-5, 8, 2)), labels=F, tck=0.01)
axis(2, pos=0, at=c(seq(-7, 6, 2)[-5]), labels=F, tck=0.01)
points(PLSDA$scores[, 1], PLSDA$scores[, 2], pch=21, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(FA$Group)], cex=1.5)

title(xlab=paste("Axis 1 (", round(att[1], 1), "%)", sep=""), line=0)
title(ylab=paste("Axis 2 (", round(att[2], 1), "%)", sep=""), line=0)

arrows(x0=0, y0=0, x1=loadings(PLSDA)[,1]*15, y1=loadings(PLSDA)[,2]*15, length=0.1, col="grey40", lty=5)

  
a=ordihull(PLSDA$scores, groups=paste(FA$SplBatch, FA$Group), lty=4, label=F, display=c("sites"), draw="polygon", col="white")
  
centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  

points(centroid, pch=21, cex=3, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1, font=2, col=c("white", "black", "black")[as.factor(var[,2])])

text(loadings(PLSDA)*16, labels=rownames(loadings(PLSDA)), col="grey40")
  
legend("topright" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1)
legend("bottomleft", legend=paste("N =", length(levels(as.factor(FA$SplBatch))),"; n =", nrow(PLSDA$scores)), bty="n", pt.cex=2)
legend("topleft", legend=expression(bold("B")), bty="n", cex=3, inset=c(-0.1,-0.05))

#dev.off()

  
```



### Regression tree

```{r}

# First version using all FA
xnam=paste0("prct$", colnames(prct))
fmla=as.formula(paste("FA$Origin ~ ", paste(xnam, collapse= "+")))

tree=rpart(fmla)

plot(tree)
text(tree)

table(predict(tree, type="class"), FA$Origin)

# Second version using ratios
newTab=round(rowSums(prct[, c(which(colnames(prct)=="C20.1w9"), which(colnames(prct)=="C22.1w9"), which(colnames(prct)=="C24.1w9"))])/prct[, -c(which(colnames(prct)=="C20.1w9"), which(colnames(prct)=="C22.1w9"), which(colnames(prct)=="C24.1w9"))], 2)

xnam=paste0("newTab$", colnames(newTab))
fmla=as.formula(paste("FA$Origin ~ ", paste(xnam, collapse= "+")))

newTree=rpart(fmla)

plot(newTree)
text(newTree)

table(predict(newTree, type="class"), FA$Origin)

#Let's see which samples are wrongly classified
rownames(allFA[which(predict(newTree, type="class")!=FA$Origin),])

```

### Regression tree

```{r}

# Let's see seasonal changes in diatoms biomarkers at the Glénan Islands site

GL=prct[which(substring(FA$Id, 7,8)=="GL"),]
date=as.POSIXct(FA[which(substring(FA$Id, 7,8)=="GL"),"SamplingDate"], format="%d/%m/%Y")
lvldate=as.POSIXct(levels(as.factor(date)), format="%Y-%m-%d")

pdf(file="FigS1.pdf", width=5, height=9)

par(las=1, mar=c(5,5,1,2), mfrow=c(3,1))

var=GL[, "C14.0"]
plot(date, var, type="n", xaxt="n", xlab="", ylab="", bty="n")

mtext(text="Proportion of FA 14:0 (%)", side=2, las=3, line=2.5, cex=1)
axis.POSIXct(side=1, x=date, at=date, format ="%b %y", labels=T, las=2)

arrows(x0=lvldate, y0=tapply(var, date, mean) + tapply(var, date, sd), x1=lvldate, y1=tapply(var, date, mean) - tapply(var, date, sd), length=0.03, angle=90, code=3)

lines(lvldate, tapply(var, date, mean))
points(lvldate, tapply(var, date, mean), pch=21, col="grey40", bg="black", cex=2)

var=GL[, "C16.1w7"]
plot(date, var, type="n", xaxt="n", xlab="", ylab="", bty="n")

mtext(text="Proportion of FA 16:1w7 (%)", side=2, las=3, line=2.5, cex=1)
axis.POSIXct(side=1, x=date, at=date, format ="%b %y", labels=T, las=2)

arrows(x0=lvldate, y0=tapply(var, date, mean) + tapply(var, date, sd), x1=lvldate, y1=tapply(var, date, mean) - tapply(var, date, sd), length=0.03, angle=90, code=3)

lines(lvldate, tapply(var, date, mean))
points(lvldate, tapply(var, date, mean), pch=21, col="grey40", bg="black", cex=2)

var=rowSums(GL[, c(which(colnames(GL)=="C15.0iso"), which(colnames(GL)=="C15.0anteiso"), which(colnames(GL)=="C16.0iso"))])
plot(date, var, type="n", xaxt="n", xlab="", ylab="", bty="n")

mtext(text="Proportion of BrFA (%)", side=2, las=3, line=2.5, cex=1)
axis.POSIXct(side=1, x=date, at=date, format ="%b %y", labels=T, las=2)

arrows(x0=lvldate, y0=tapply(var, date, mean) + tapply(var, date, sd), x1=lvldate, y1=tapply(var, date, mean) - tapply(var, date, sd), length=0.03, angle=90, code=3)

lines(lvldate, tapply(var, date, mean))
points(lvldate, tapply(var, date, mean), pch=21, col="grey40", bg="black", cex=2)

dev.off()


```



## Stable isotopes
### Graphical representation (biplot 13C-15N)

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

#cairo_pdf(file="Fig3_SI.Biplot.pdf", width=7.5, height=6)

par(las=1, mar=c(5,5,1,2), bty="n", mfrow=c(1,1))

plot(SI$d13C, SI$d15N, bty="n", xlim=c(-20,-12), ylim=c(6, 16), xlab=expression(paste(delta^13, "C (\u2030)")), ylab=expression(paste(delta^15, "N (\u2030)", )), type="n")

axis(1, pos=0, at=c(seq(-20, 12, 4)), labels=F, tck=0.01)
axis(2, pos=0, at=c(seq(6, 16, 2)), labels=F, tck=0.01)
points(SI$d13C, SI$d15N, pch=21, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(SI$Group)], cex=1.5)
  
a=ordihull(cbind(SI$d13C, SI$d15N), groups=paste(SI$SplBatch, SI$Group), lty=4, label=F, display=c("sites"), draw="polygon", col="white")
  
centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  

points(centroid, pch=21, cex=3, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1, font=2, col=c("white", "black", "black")[as.factor(var[,2])])

legend("topleft" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1, inset=c(0.02, 0))
legend("bottomleft", legend=paste("N =", length(levels(as.factor(SI$SplBatch))),"; n =", nrow(SI)), bty="n", pt.cex=2)

#dev.off()
  
```

## Amino acids

```{r}
allAA=AA[,-c(1:5)]
rownames(allAA)=AA[,1]

# Let's edit a summary table

summaryAA=data.frame(M=aggregate(allAA , by=list(AA$SplBatch), mean), SD=aggregate(allAA, by=list(AA$SplBatch), sd), LE=aggregate(allAA, by=list(AA$SplBatch), length))

#write.table(t(summaryAA), file="meansAA.txt")

```

### Graphical representation using PCA

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

#svg(file="Fig4A.svg", width=7.5, height=6)

par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

acp=dudi.pca(allAA, scannf = FALSE, nf = 2)
varPCA=get_pca_var(acp)
varPCAhigh=which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*1.2|varPCA$contrib[,2]>100/nrow(varPCA$contrib)*1.2)
varPCAlow=which(varPCA$contrib[,1]<100/nrow(varPCA$contrib)*1.2&varPCA$contrib[,2]<100/nrow(varPCA$contrib)*1.2)

plot(acp$li[,c(1,2)], acp$co[,c(1,2)], type="n", ann=F, bty="n", axes=F, ylim=c(-5, 4), xlim=c(-6, 6))
axis(1, pos=0, at=c(seq(-6, 6, 2)[-4]), labels=F, tck=0.01, padj=-4)
axis(2, pos=0, at=c(seq(-5, 4, 2)), labels=F, tck=0.01, hadj=0.1)

title(xlab=paste("PC1 (", round((acp$eig/sum(acp$eig)*100)[1], 1), "%)", sep=""), line=0)
title(ylab=paste("PC2 (", round((acp$eig/sum(acp$eig)*100)[2], 1), "%)", sep=""), line=0)

arrows(x0=0, y0=0, x1=acp$co[,1][varPCAhigh]*5, y1=acp$co[,2][varPCAhigh]*5, length=0.1, col="grey40", lty=5)
text(acp$co[varPCAhigh,]*5.5, labels=colnames(allAA)[varPCAhigh], col="grey40")

arrows(x0=0, y0=0, x1=acp$co[,1][varPCAlow]*5, y1=acp$co[,2][varPCAlow]*5, length=0.1, col="grey90", lty=5)
text(acp$co[varPCAlow,]*5.5, labels=colnames(allAA)[varPCAlow], col="grey90")

a=ordihull(acp$li, groups=paste(AA$SplBatch, AA$Group), lty=4, label=F, display=c("sites"), draw="lines")

points(acp$li[, 1], acp$li[, 2], pch=23, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(AA$Group)], cex=1.5)

a=ordihull(acp$li, groups=paste(AA$SplBatch, AA$Group), border=NA, label=F, display=c("sites"), draw="polygon", col="white", alpha=50)

centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  

points(centroid, pch=21, cex=3.5, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1.2, font=2, col=c("white", "black", "black")[as.factor(var[,2])])

legend("topright" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1)

legend("bottomleft", legend=paste("N =", length(levels(as.factor(AA$SplBatch))),"; n =", nrow(acp$li)), bty="n", pt.cex=2)
legend("topleft", legend=expression(bold("A")), bty="n", cex=3, inset=c(-0.1,-0.05))


#dev.off()
  
```


### Graphical representation using PLSDA

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

# Transforming group into indicating variable
var.ind <- DUMMY(paste(AA$Group, sep=""))
fac <- as.factor(paste(AA$Group, sep=""))

# Put the matrix into the good "matrix" shaping
DATA.M <- as.matrix(allAA)

# Canonical Powered Partial least squared (PLS) regression
PLSDA<-cppls(var.ind ~ DATA.M, ncomp=2, scale=T) 
summary (PLSDA)
attributes(PLSDA)
PLSDA$coefficients
att=attributes(loadings(PLSDA))$explvar

# Drawing

#svg(file="Fig4B_AA.PLSDAbis.svg", width=7.5, height=6)

par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

plot(PLSDA$scores, type="n", ann=F, bty="n", axes=F, xlim=c(-6, 6), ylim=c(-2, 3))
  
axis(1, pos=0, at=c(seq(-6, 6, 2)[-4]), labels=F, tck=0.01)
axis(2, pos=0, at=c(seq(-2, 3, 2)), labels=F, tck=0.01)
points(PLSDA$scores[, 1], PLSDA$scores[, 2], pch=21, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(AA$Group)], cex=1.5)

title(xlab=paste("Axis 1 (", round(att[1], 1), "%)", sep=""), line=0)
title(ylab=paste("Axis 2 (", round(att[2], 1), "%)", sep=""), line=0)

arrows(x0=0, y0=0, x1=loadings(PLSDA)[,1]*3.5, y1=loadings(PLSDA)[,2]*3.5, length=0.1, col="grey40", lty=5)
text(loadings(PLSDA)*3.6, labels=rownames(loadings(PLSDA)), col="grey40")

  
a=ordihull(PLSDA$scores, groups=paste(AA$SplBatch, AA$Group), lty=4, label=F, display=c("sites"), draw="polygon", col="white")
  
centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  


points(centroid, pch=21, cex=3, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1, font=2, col=c("white", "black", "black")[as.factor(var[,2])])
  
  
legend("topright" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1)
legend("bottomleft", legend=paste("N =", length(levels(as.factor(AA$SplBatch))),"; n =", nrow(PLSDA$scores)), bty="n", pt.cex=2)
legend("topleft", legend=expression(bold("B")), bty="n", cex=3, inset=c(-0.1,-0.05))


#dev.off()

  
```

### Regression tree

```{r}

# First version using all AA
xnam=paste0("allAA$", colnames(allAA))
fmla=as.formula(paste("AA$Origin ~ ", paste(xnam, collapse= "+")))

tree=rpart(fmla)

plot(tree)
text(tree)

table(predict(tree, type="class"), AA$Origin)

# Second version using ratios
newTab=round(allAA[, which(colnames(allAA)=="Asx")]/allAA[, -which(colnames(allAA)=="Asx")], 2)

xnam=paste0("newTab$", colnames(newTab))
fmla=as.formula(paste("AA$Origin ~ ", paste(xnam, collapse= "+")))

newTree=rpart(fmla)

plot(newTree)
text(newTree)

table(predict(newTree, type="class"), AA$Origin)

#Let's see which samples are wrongly classified
rownames(allAA[which(predict(newTree, type="class")!=AA$Origin),])

```

## Sterols

```{r}
allSt=St[,-c(1:6, which(colnames(St)=="Stigmasterol"))]
rownames(allSt)=St[,1]
allSt[is.na(allSt)]=0

## Percentage table
prctSt=allSt/rowSums(allSt)*100

## Concentrations table 
concSt=allSt/St[,which(colnames(St)=="Stigmasterol")]*St[,which(colnames(St)=="StdStig")]/St[,which(colnames(St)=="SplMass")]
conctotSt=rowSums(concSt)

# Let's edit a summary table

summarySt=data.frame(M=aggregate(cbind(prctSt, conctotSt) , by=list(St$SplBatch), mean), SD=aggregate(cbind(prctSt, conctotSt), by=list(St$SplBatch), sd), LE=aggregate(cbind(prctSt, conctotSt), by=list(St$SplBatch), length))

#write.table(t(summarySt), file="meansSt.txt")

```

### Graphical representation using PCA

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

#svg(file="Fig5A_PCA.svg", width=7.5, height=6)

par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

acp=dudi.pca(sqrt(prctSt), scannf = FALSE, nf = 2, scale=T)
varPCA=get_pca_var(acp)
varPCAhigh=which(varPCA$contrib[,1]>100/nrow(varPCA$contrib)*1.2|varPCA$contrib[,2]>100/nrow(varPCA$contrib)*1.2)
varPCAlow=which(varPCA$contrib[,1]<100/nrow(varPCA$contrib)*1.2&varPCA$contrib[,2]<100/nrow(varPCA$contrib)*1.2)

plot(acp$li[,c(1,2)], acp$co[,c(1,2)], type="n", ann=F, bty="n", axes=F, xlim=c(-6, 8), ylim=c(-6, 5))
axis(1, pos=0, at=c(seq(-6, 8, 2)[-4]), labels=F, tck=0.01, padj=-4)
axis(2, pos=0, at=c(seq(-6, 5, 2)[-4]), labels=F, tck=0.01, hadj=0.1)

title(xlab=paste("PC1 (", round((acp$eig/sum(acp$eig)*100)[1], 1), "%)", sep=""), line=0)
title(ylab=paste("PC2 (", round((acp$eig/sum(acp$eig)*100)[2], 1), "%)", sep=""), line=0)

arrows(x0=0, y0=0, x1=acp$co[,1][varPCAhigh]*4.5, y1=acp$co[,2][varPCAhigh]*4.5, length=0.1, col="grey40", lty=5)
text(acp$co[varPCAhigh,]*5, labels=colnames(prctSt)[varPCAhigh], col="grey40")

arrows(x0=0, y0=0, x1=acp$co[,1][varPCAlow]*5, y1=acp$co[,2][varPCAlow]*5, length=0.1, col="grey90", lty=5)
text(acp$co[varPCAlow,]*5.5, labels=colnames(prctSt)[varPCAlow], col="grey90")

a=ordihull(acp$li, groups=paste(St$SplBatch, St$Group), lty=4, label=F, display=c("sites"), draw="lines")

points(acp$li[, 1], acp$li[, 2], pch=23, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(St$Group)], cex=1.5)

a=ordihull(acp$li, groups=paste(St$SplBatch, St$Group), border=NA, label=F, display=c("sites"), draw="polygon", col="white", alpha=50)

centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  

points(centroid, pch=21, cex=3.5, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1.2, font=2, col=c("white", "black", "black")[as.factor(var[,2])])

legend("topright" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1)

legend("bottomleft", legend=paste("N =", length(levels(as.factor(St$SplBatch))),"; n =", nrow(acp$li)), bty="n", pt.cex=2)
legend("topleft", legend=expression(bold("A")), bty="n", cex=3, inset=c(-0.1,-0.05))


#dev.off()
  
```


### Graphical representation using PLSDA

```{r echo=FALSE, fig.width=7.5, fig.height=6, fig.align="center"}

# Transforming group into indicating variable
var.ind <- DUMMY(paste(St$Group, sep=""))
fac <- as.factor(paste(St$Group, sep=""))

# Put the matrix into the good "matrix" shaping
DATA.M <- as.matrix(prctSt)

# Canonical Powered Partial least squared (PLS) regression
PLSDA<-cppls(var.ind ~ DATA.M, ncomp=2, scale=T) 
summary (PLSDA)
attributes(PLSDA)
PLSDA$coefficients
att=attributes(loadings(PLSDA))$explvar


# Drawing

#svg(file="Fig5B_PLSDA.svg", width=7.5, height=6)

par(las=1, mar=c(2,2,1,2), bty="n", mfrow=c(1,1))

plot(PLSDA$scores, type="n", ann=F, bty="n", axes=F, , xlim=c(-4, 7), ylim=c(-4, 4))

axis(1, pos=0, at=c(seq(-4, 7, 2)[-3]), labels=F, tck=0.01)
axis(2, pos=0, at=c(seq(-4, 4, 2)[-3]), labels=F, tck=0.01)
points(PLSDA$scores[, 1], PLSDA$scores[, 2], pch=21, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(St$Group)], cex=1.5)
  
title(xlab=paste("Axis 1 (", round(att[1], 1), "%)", sep=""), line=0)
title(ylab=paste("Axis 2 (", round(att[2], 1), "%)", sep=""), line=0)

arrows(x0=0, y0=0, x1=loadings(PLSDA)[,1]*7, y1=loadings(PLSDA)[,2]*7, length=0.1, col="grey40", lty=5)

text(loadings(PLSDA)*7, labels=rownames(loadings(PLSDA)), col="grey40")

a=ordihull(PLSDA$scores, groups=paste(St$SplBatch, St$Group), lty=4, label=F, display=c("sites"), draw="polygon", col="white")
  
centroid=t(summary(a))
var=data.frame(do.call("rbind", strsplit(rownames(centroid), " ", fixed=F)))  

points(centroid, pch=21, cex=3, col="grey40", bg=c("grey20", "grey80", "white")[as.factor(var[,2])])
text(centroid, labels=var[,1], cex=1, font=2, col=c("white", "black", "black")[as.factor(var[,2])])
  
  
legend("topright" , legend=c("Wild", "Captured and caged", "Born in captivity"), bty="n", pch=21, pt.bg=c("white", "grey80", "grey20"), col="grey40", pt.cex=2, cex=1)
legend("bottomleft", legend=paste("N =", length(levels(as.factor(St$SplBatch))),"; n =", nrow(PLSDA$scores)), bty="n", pt.cex=2)
legend("topleft", legend=expression(bold("B")), bty="n", cex=3, inset=c(-0.1,-0.05))


#dev.off()

  
```

### Regression tree

```{r}

# First version using all sterols
xnam=paste0("prctSt$", colnames(prctSt))
fmla=as.formula(paste("St$Origin ~ ", paste(xnam, collapse= "+")))

tree=rpart(fmla)

plot(tree)
text(tree)

table(predict(tree, type="class"), St$Origin)

# Second version using ratios
newTab=round(prctSt[, which(colnames(prctSt)=="C29H46")]/prctSt[, -which(colnames(prctSt)=="C29H46")], 2)

xnam=paste0("newTab$", colnames(newTab))
fmla=as.formula(paste("St$Origin ~ ", paste(xnam, collapse= "+")))

newTree=rpart(fmla)

plot(newTree)
text(newTree)

table(predict(newTree, type="class"), St$Origin)

#Let's see which samples are wrongly classified
rownames(allSt[which(predict(newTree, type="class")!=St$Origin),])

```





